---
title: 'Additive Design: The Concept and Data Analysis'
author: "Maxwel C. Oliveira, Gustavo A. M. Pereira, Rodrigo Werle, and Stevan Z. Knezevic"
date: "January 24, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

##Download R and R-Studio 
At <https://www.r-project.org/>, go to download in the left column, click in CRAN and choose a location near you. For example, in Brazil you can choose seven different locations. Then download R compatible to your system (Linux, Mac or Windows). Once you have downloaded **R**, you should also download **R-Studio** <https://www.rstudio.com/>. **R-Studio** is a friendly interface for R programers. Scroll down to **R-Studio** and click in Download. Choose the free version of **R-Studio**, there will also  be **R-Studio** version for Linux, Mac, and Windows. 


##Create an R-studio file
Open **R-Studio** at the toolbar, click in file, New Project…, Existing Directory, and choose the folder that contain your data. Your data file have to be in that chosen folder. In addition, we recommend your data to be saved as csv (comma delimited) file.

##Load the data in R-studio
Assign the name of your data set (replace “DMT” to a name of your choice). If you use comma (,) for separating decimals places use *read.csv2()*. If you use a period (.), use *read.csv()*, as described above. In parenthesis, write the name of your data set file. 

```{r LoadData, warning=F, include=TRUE}
DMT=read.csv("dmshoot.csv")
```


The command *head* prints the first six lines of the data set. It is useful for double checking your data set.
```{r head, warning=F, include=TRUE}
head(DMT)
```


The command *str* prints how **Rstudio** is reading the characters vector in a data set.
```{r str, warning=F, include=TRUE}
str(DMT)
```

## Cousens Model

The empirical model: *$Y = I * x  /  1 + (I/A)*x$* is the standard model to describe additive competition studies. *I* reoresents the slope of *Y* (yield loss) when *x* (weed density) aproximate zero. In addition, *A* is the assymtopte or maximum expected yield loss (%).


##Fit a full model, Cousens model with 4 parameters (Step 1).

**Full** is a user-defined name that will contain all information about the fitted model generated by nls (nonlinear least squares) function. The *start* is used to estimate values of parameter *I* and *A* for the model. Parameters can easier determined from visual inspection of the data set (plotting data and observing trends). The bracket [] after each parameter in the equation tell **R** to estimate a parameter for each weed species (4 parameters).
```{r Full model, warning=F}
Full = nls(yl ~ (I[weed]*densityweed)/(1+(I[weed]/A[weed])*densityweed), data=DMT, start=list(I=c(60,30), A=c(80,60)), trace=T)
```

##Check estimated parameters.
The *summary* command provides the estimated parameters *I* and *A* for each weed species, *Commelina benghalensis* (species 1) and *Richardia brasiliensis* (species 2) .
```{r Fullparm, warning=F}
summary(Full)
```


##Fit a reduced model (Red.1), Cousens model with 2 parameters (Step 2).
**Red.1** is a user-defined name that will contain information about the first reduced model generated by the nls function. Notice that we do not include bracket [] after each parameter *I* and *A*. In this case we are combining parameter *I* and *A* for both weed species. Our hypothesis is that a single parameter *I* and *A* for both species is enough to describe the crop-weed relationship. 
```{r Red.1 model, warning=F}
Red.1 = nls(yl ~ (I*densityweed)/(1+(I/A)*densityweed), data=DMT, 
start=list(I=40, A=80), trace=T)
```

##Check estimated parameters.
This command provides the estimated parameters *I* and *A* for both weed species together.
```{r Red.1parm, warning=F}
summary(Red.1)
```


##Test a first hypothesis
Testing hypothesis using *ANOVA*. We estimated using the **Full** model (*I* and A for each species) to compare with **Red.1** (single *I* and *A* for both species). If P-value>0.05, models are similar; therefore we should use the **Red.1** model, which means that the simplest model (**Red.1**) is appropriate to describe crop-weed relationship. If not we should proceed to the next hypothesis.
```{r hyp1, warning=F}
anova(Full, Red.1)
```
The F-test showed P<0.05, therefore the Red.1 model is not appropriate to describe the crop-weed relationship.


##Fit a reduced model (Red.2), Cousens model with 3 parameters (Step 3).
**Red.2** is a user-defined name that will contain information about the second reduced model generated by *nls* function. Notice that the bracket [] is after the parameter *A* only, which means that we are testing a hypothesis of single parameter *I*, but different *A* for the species. 
```{r Red.2 model, warning=F}
Red.2 = nls(yl ~ (I*densityweed)/(1+(I/A[weed])*densityweed), data=DMT, start=list(I=60, A=c(80,60)),  trace=T)
```

##Check estimated parameters.
This command provides the estimated parameters *I* for both weed species and *A* for each weed species.
```{r Red.2parm, warning=F}
summary(Red.2)
```

##Test a second hypothesis
Testing hypothesis using F-test. We are using the Full model (separated I and *A* for each species) to compare with Red.2 (single *I* and different *A* for both species). If P-value>0.05, models are similar; therefore, we should use the **Red.2** model, which means that the simplest model (**Red.2**) is appropriate to describe crop-weed relationship. If not we should proceed to the next hypothesis.
```{r hyp2, echo=T, warning=F}
anova(Full, Red.2)
```
The F-test showed P<0.05, therefore the Red.2 model is not appropriate to describe the crop-weed relationship.



##Fit a reduced model (Red.3), 3 parameters (Step 4).
Red.3 is a user-defined name that will contain information about the third reduced model generated by *nls* function. Notice that the bracket [] is after the parameter *I* only, which means that we are testing a hypothesis of different parameter *I*, but single parameter *A* for the species.
```{r Red.3 model, warning=F}
Red.3 = nls(yl ~ (I[weed]*densityweed)/(1+(I[weed]/A)*densityweed), data=DMT, start=list(I=c(30,30), A=70), trace=T)
```

##Check estimated parameters.
This command provides the estimated parameters *I* for each weed species and *A* for both weed species.
```{r Red.3parm, warning=F}
summary(Red.3)
```

##Test a third hypothesis
Testing hypothesis using F-test. We are using the Full model (separated *I* and *A* for each species) to compare with Red.3 (different *I* and single *A* for both species). If P-value>0.05, models are similar; therefore we should use the **Red.3** model, which means that the simplest model (**Red.3**) is appropriate to describe crop-weed relationship.
```{r hyp3, warning=F}
anova(Full, Red.3)
```
Results showed that P >0.05. Therefore, the **Full** model can be simplified to **Red.3** model.

##Plotting the Red.3 model
The command *par* is used to define the plot size.
The command *plot* and *lines* are used to generate the figure and the averaged points of yield loss at each density (Figure 4). The command *subset* is adding each weed species separately in the plot (weed 1) and lines (weed 2). 

The x is a user-defined name, it will contain the x-axis sequence of the data set. weed1 and weed2 is also a user-defined name, this is the equation with the previous parameter estimates *I* and *A* estimated from Red.3 model using the nls function. Notice that the parameters estimated in Red.3 model were inserted in the Cousens model for each weed species.


The command *lines* will insert the previous equation into the plot. Command *lty*, *lwd*, and *col* define the line type, size, and color, respectively.

The command *legend* will add the legend into the plot area.
```{r plot, warning=F}
par(mar=c(5,6,2,2), mgp=c(3,1.5,0))
plot(yl~densityweed, data=DMT, subset = weed =="1", pch=16, cex=1, las=1, xlab=expression("Weed Density (plants pot"^-1*")"), ylim=c(-10,110), ylab = "Yield Loss (%)", cex.axis=1, cex.lab=1)
lines(yl~densityweed, type="p",data=DMT, subset = weed =="2", col=2, cex=1, pch=1)

x=seq(0,4,0.25)
weed1=(228.357*x)/(1+(228.357/106.170)*x)
weed2=(37.000*x)/(1+(37.000/106.170)*x)

lines(x,weed1, lty=1, lwd=1, col=1)
lines(x,weed2, lty=3, lwd=1, col=2)

legend("bottomright", legend=c("C. benghalensis", "R. brasiliensis"), text.font = 3, col=c(1,2), pch= c(16,1), lty=c(1,3), lwd= c(1,1), bty="n", cex=1)
```




A good package for producing high quality figure in **R** is the *ggplot2*. 

```{r library, warning=F}
library(ggplot2)
library(broom)
```


```{r Red model, warning=F}
DMT$weed<-factor(DMT$weed, levels=c("1", "2"), labels=c("Commelina benghalensis", "Richadia brasiliensis"))


Red.3 = nls(yl ~ (I[weed]*densityweed)/(1+(I[weed]/A)*densityweed), data=DMT, alg="port", 
                      start=list(I=c(30,30), A=70), trace=T)
summary(Red.3)



nd1 = data.frame(densityweed=seq(0, 4, 0.01), weed="Commelina benghalensis")
nd2 = data.frame(densityweed=seq(0, 4, 0.01), weed="Richadia brasiliensis")
nd = rbind(nd1, nd2)

pred<- augment(Red.3, newdata=nd)
```

```{r Red model plot}
ggplot(DMT, aes(x=densityweed, y=yl, color=weed)) + geom_point(shape=1, size=3) + geom_line(data = pred, size=1.3, aes(x=densityweed, linetype=weed, y=.fitted)) +
  labs(fill="", y="Yield loss (%)", x=expression(bold(paste("Weed density (plants m"^"-2",")")))) +
  scale_colour_manual(values = c("red", "black"))+
  scale_y_continuous(limits=c(-25,110), breaks = c(-25,0,25,50,75,100)) +
    theme(axis.text=element_text(size=15, color="black"), 
        axis.title=element_text(size=17,face="bold"), 
        panel.background = element_rect(fill="white", color = "white"), 
        panel.grid.major = element_line(color = "white"), 
  panel.grid.minor = element_blank(), panel.border = element_rect(fill=NA,color="black", size=0.5, 
  linetype="solid"), legend.position=c(0.7,0.15), legend.text = element_text(size = 12, colour = "black", face="italic"), legend.key = element_rect(fill=NA), legend.key.height  = unit(1.5, "line"), legend.key.width = unit(2.2, "line"),
legend.background = element_rect(fill =NA),  legend.title=element_blank())  +  ggsave("Red.tiff", units="in", width=6, height=6, dpi=300)
```



## AICc model selection and Goodness of fit

According to the AICc criterion, the top model has the lowest AICc value. The AICc calculation can be simplified using **R**, the first step is loading the package *AICcmodavg*.

```{r aic library}
library(AICcmodavg)
```


The four candidate models using the rectangular hyperbola are compared using AICc.
```{r aic table, warning=FALSE}
cand.mods<- list(Full, Red.1, Red.2, Red.3)
 
Modnames<- c('Full',' Red.1',' Red.2',' Red.3')

aictab(cand.set = cand.mods,modnames = Modnames, sort = TRUE)
```


Root mean square error (RMSE) for goodness of fit of the top model (**Red.3**) selected.
```{r assessing fit of top model, warning=F}
mse <- mean(residuals(Red.3)^2)
rmse <- sqrt(mse)
rmse
```




